# Scale

In geom, `aes(xxx=yyy)`, defines a mapping from data vector `yyy` to aesthetic vector `xxx`. To change or fine tune this mapping, you add `scale_xxx_...` to your ggplot object.

In general, a mapping can be divided into two parts

  * How to map?  
    * `limits`: the range of data vector to be mapped.  
    * `values`: the mapped values from `limits`.  
    
  * In the the mapping, what to show to the readers? This is about the information in **legend** (圖例說明)  
    * `breaks`: the picked values in `limits`.  
    * `labels`: the mapped values of `breaks`.  
    * `name`: the title of the legend.

Depending on aesthetics, not all five settings will be necessary. 

### `scale_x_datetime`/`scale_x_date`

```{r}
scale_x_date()-> xx
```

***

```{r etef}
teachDS::img_centering("https://www.economist.com/img/b/1280/748/90/sites/default/files/20190112_WOC124.png", width="80%")
```

```{r}
dataSet1 <- 
  data.frame(
    x=1979:2018
  )
dataSet1$y <- sample(10:40, length(dataSet1$x), T)
ggplot1 <- list()

ggplot()+
  geom_step(
    data=dataSet1,
    mapping=
      aes(
        x=x,
        y=y
      )
  ) -> ggplot1$plot0
ggplot1$plot0
```

```{r}
ggplot1$plot0 +
  scale_x_continuous(
    breaks=c(1979, 1985, 1990, 1995, 2000, 2005, 2010, 2015, 2018),
    labels=c("1979", "85", "90", "95", "2000", "05", "10", "15", "18")
  ) -> ggplot1$plot1

ggplot1$plot1
```

```{r}
ggplot1$plot1 +
  geom_rug(
    mapping=aes(
      x=1979:2018
    ),
    outside=TRUE,
    size=1, #input$size
    length=grid::unit(
      0.1, #input$length
      "mm"
    )
  ) +
  coord_cartesian(clip="off")

```

